Figures and Plots

### 2025-06-17 docs/manuscript_GNM.qmd
Some sensitivity analysis added

@@

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details the study's objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

@@

While lnRR and SMDH differ in their measurement scale, both effect sizes suggest a small positive but non-conclusive effect of green nest material on fitness. These differences in overall effect size magnitude likely reflect the underlying biological and statistical assumptions - lnRR incorporates the mean difference at a multiplicative scale whereas SMDH expresses the difference on an additive scale. The difference in relative heterogeneity (I^2^) between the two effect sizes aligns with broader trends observed in meta-analysis in ecology and evolution (@yang2024). As a pre-registered sensitivity analysis, we fitted a bivariate multilevel meta-analytic model including both the effect sizes. While neither intercept was statistically significant, the results strengthens the interpretation that any effect of green nest material on fitness outcomes in likely to be very small at best \[lnRR = `{r} as.double(formatC(coef(bivariate_model)[1], digit = 3, format="f"))`, p-value = `{r} as.double(formatC(bivariate_model$pval[1], digits = 3, format = "f"))` and SMDH = `{r} as.double(formatC(coef(bivariate_model)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(bivariate_model$pval[2], digits = 3, format = "f"))`\].

@@

Most bird species, showed a small positive non-significant increase it fitness with an experimental increase in green nest material, although some species showed negative but similarly non-significant effects \[[\@figsupp-MR](C-D); Table X\]. All confidence and prediction intervals overlapped zero. However, bird species identity explained a proportion of heterogeneity in the effect sizes ($R^2_{marginal(lnRR)}$= `{r} as.double(round(orchaRd::r2_ml(lnRR_birds)*100,1))[1]`% and $R^2_{marginal(SMDH)}$= `{r} as.double(round(orchaRd::r2_ml(SMDH_birds)*100,1))[1]`%).

@@

>>>>

# Introduction

Nest construction is a fundamental aspect of the bird breeding cycle, providing structural support and protection for eggs, parents, and nestlings against predators and environmental pressures, and in some species, even playing a role in sexual display during mating [@mainwaring2014; @hansell2000]. Nest building behaviour is increasingly recognized as an expression of extended phenotype, demonstrating interspecific and intraspecific trait variation [@moreno2012; @perez2023]. While most birds use organic plant materials- twigs, branches, grass, and moss- some incorporate inorganic items such as stones, feathers, snake skin and even anthropogenic materials [@sheard2024]. These non-structural additions suggest that nests are more than just physical shelters; they can be complex tools shaped by both natural and sexual selection for multiple functions.

An intriguing nesting behaviour observed in some birds is the use of fresh green plant material (green nest material or GNM). This includes leaves, sprigs, herbs and branches added during or even after primary nest construction has completed. Importantly, these green nest materials differ from the structural components, being selected from a small, non-random subset of local flora [@clark1985; @garrido-bautista2023], often characterized by high levels of aromatic volatile compounds (commonly used by Passerines, [@mennerat2009]) and secondary metabolites such as oils and resins (preferred by raptors, [@ontiveros2007]). The costs associated with gathering these materials beyond the territory and maintaining fresh greenery through the breeding season imply adaptive significance of the GNM.

Several hypotheses have been proposed to explain the adaptive function of green nest material (GNM). The courtship hypothesis suggests that males use aromatic plants as olfactory signals to attract mates or advertise nest site occupancy, particularly in cavity-nesting species where visual cues are limited. This idea traces back to the early natural history observations, beginning with @kessel1957, who noted that male European Starlings (*Sturnus vulgaris*) carried fresh plant material to nests while singing and displaying to females. Subsequent studies in European and Spotless Starlings (*Sturnus unicolor*) support this view: unpaired males conspicuously deliver greenery during courtship but stop after egg-laying begins [@brouwer2004; @veiga2006], and they increase this behaviour in the presence of potential mates \[[@eens1993]\]. Females often prefer nests with fresh herbs, suggesting that GNM may serve as a sexually selected signal of male quality and potentially increase fitness benefits for females. The courtship hypothesis has been also extended to species like Blue Tits (*Cyanistes caeruleus*), where it has been proposed that GNM might serve as a signal from females to the males [@tomás2013]. However, in both starlings and blue tits, the link between plant-bringing material in males and the fitness benefits for females still remain unclear.

Another hypothesis to explain the adaptive benefit of GNM is the widely tested nest protection hypothesis (NPH), which proposes that volatile compounds in green material reduce parasite abundance or pathogen loads, indirectly enhancing offspring survival. In an influential early comparitive study in raptors (Falconiformes), @wimberger1984 found indirect evidence for the reduction of ectoparasites in the nests with green nest material. @clark1988 later provided experimental support in European Starlings, showing that preferred herbs contained insect-repellent and antimicrobial volatile. They also outlined key criteria for evaluation of the NPH, including selective plant use, bioactive chemical properties and demonstrable adaptive benefits. Field observation and experimental studies on raptors [@ontiveros2007; @dykstra2009] and passerines [@dawson2004; @lafuma2001; @mennerat2009b; @quiroga2012] have shown mixed results, and in some cases, GNM improved nestling condition without reducing parasite abundance [@gwinner2000]. These findings led to the development of drug hypothesis, which suggests that the aromatic GNM directly enhance nestling development or immune function, independent of the parasite control [@gwinner2000; @mennerat2009a]. Thus, while both hypotheses posit adaptive benefits of GNM, the drug hypothesis emphasizes a pharmacological mechanism, and the two are not mutually exclusive. Several other adaptive functions of GNM have also been proposed. These functions do not refer to the aromatic properties of GM, but could still provide some benefits such as enhance the crypticity of the nest [@skutch1957], provide insulation for eggs and nestling [@mertens1977], provide protection from sun, signal nest occupancy or simply fill the gaps between nesting materials [@lyons1987]. These alternative functional roles have received little support or attention.

Despite several observational and experimental studies and two crucial narrative reviews [@dubiec2013; @scott-baumann2015], a comprehensive quantitative synthesis across studies and hypotheses has been lacking. This has limited our understanding of the generality and magnitude of GNM effects on fitness benefits. To address this gap, we conducted a systematic review and meta-analysis to assess the overall effect of GNM on various fitness proxies. We explicitly evaluate the evidence for the courtship and parental-care- the latter combining the nest protection and drug hypotheses due to their non-exclusivity. The drug hypothesis can only be supported by improved nestling health while ensuring that GNM has no protective effect against any parasite or pathogen.

In addition to testing these primary hypotheses, we further explore the context-dependency in the research field and examine the potential sources of variation in the results (heterogeneity) in the GNM effects. Specifically, we explored whether the effect of GNM varied depending on the type of parasite (arthropods vs. microorganisms), the timing of GNM addition (e.g., before hatching, after hatching, or continuously), the experimental design (e.g., aromatic vs. non-aromatic vs. no added material), the focal species, and the type of fitness trait measured (e.g., physiology, morphology, reproduction). These exploratory moderators could allow us to assess context dependence and identify ecological or methodological patterns that may explain inconsistencies in previous findings. Together, our evidence synthesis approach provides the first quantitative test of the adaptive value of green nest material across bird species and research contexts.

@@

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details our objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

@@

As a pre-registered sensitivity analysis, we fitted a bivariate multilevel meta-analytic model including both the effect sizes. While neither intercept was statistically significant, the results strengthens the interpretation that any effect of green nest material on fitness outcomes in likely to be very small at best \[lnRR = `{r} as.double(formatC(coef(bivariate_model)[1], digit = 3, format="f"))`, p-value = `{r} as.double(formatC(bivariate_model$pval[1], digits = 3, format = "f"))` and SMDH = `{r} as.double(formatC(coef(bivariate_model)[2], digit = 3, format="f"))` , p-value = `{r} as.double(formatC(bivariate_model$pval[2], digits = 3, format = "f"))`\].

@@

Most bird species, showed a small positive non-significant increase it fitness with an experimental increase in green nest material, although some species showed negative but similarly non-significant effects \[[\@figsupp-MR](C-D); Table X\]. However, bird species identity explained a proportion of heterogeneity in the effect sizes ($R^2_{marginal(lnRR)}$= `{r} as.double(round(orchaRd::r2_ml(lnRR_birds)*100,1))[1]`% and $R^2_{marginal(SMDH)}$= `{r} as.double(round(orchaRd::r2_ml(SMDH_birds)*100,1))[1]`%).

@@

# Discussion

Our systematic review and meta-analysis found a small positive increase in fitness benefits in the bird nests with increased green nest material, though the strength of these effects depended on the effect size metric used. SMDH showed a larger effect with the There is high heterogeneity among the effect sizes, emphasizing the context-dependence of this benefit. Importantly, our results reveal that the commonly studied mechanistic hypotheses of green nest material---whether through sexual signaling, parasite reduction, or direct nestling benefits---lack consistent empirical support across studies.

The While lnRR and SMDH differ in their measurement scale, both effect sizes suggest a small positive but non-conclusive effect of green nest material on fitness. These differences in overall effect size magnitude likely reflect the underlying biological and statistical assumptions - lnRR incorporates the mean difference at a multiplicative scale whereas SMDH expresses the difference on an additive scale. The difference in relative heterogeneity (I^2^) between the two effect sizes aligns with broader trends observed in meta-analysis in ecology and evolution (@yang2024).

### 2025-06-17 docs/notebook/data_analysis.qmd
Some sensitivity analysis added

@@

@@

@@

```{r Types of parasite}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_lnRR_parasite <- dataset_lnRR%>%
  filter(!is.na(parasite_type))

@@

## Heterogeneity

To understand the sources of variation or heterogeneity,

### For the intercept-only models

we will use the pluralistic approach recently suggested by Yang et al. (2023),

which consists of calculating four metrics of heterogeneity (σ2, I2, CV, M) to comprehensively understand both total and random-effect specific heterogeneity.

### For the meta-regressions,

we will calculate Rmarginal2 to quantify the heterogeneity explained by the moderator(s) in proportion to the total variation (Nakagawa & Schielzeth, 2013).

@@

To explore evidence of publication bias, we will run multilevel meta-regressions with the same random effects structure as described above. We will follow the approach recommended in section 4.3 from Nakagawa et al. (2022) for both lnRR and SMDH, which consists of the following steps:

First, run a uni-moderator meta-regression to test for evidence of small-study effects for both lnRR and SMDH separately, where the moderator is the square root of the inverse of the 'effective sample size' (see equation 27 in Nakagawa et al. 2022). The slope of this meta-regression provides information about whether funnel plot asymmetry exists, and the intercept corresponds to the overall effect size adjusted by small-study effects. However, if this intercept is statistically significantly different from zero,

```{r}
dataset_SMDH<-dataset_SMDH%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))

MR_SMDH_small_study <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = SMDH_variance,# specify lnRR's sampling variance;
                    mods = ~sqrt_effective_n_inv,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

summary(MR_SMDH_small_study)

SMDH_small_study_bubble<-orchaRd::mod_results(MR_SMDH_small_study,
                     mod="sqrt_effective_n_inv",
                     group="paper_ID",
                     weights="prop")
orchaRd::bubble_plot(SMDH_small_study_bubble,
                      mod="sqrt_effective_n_inv",
                     group="paper_ID")

round(r2_ml(MR_SMDH_small_study)*100,2)
```

Then, run the following uni-moderator meta-regression for both lnRR and SMDH separately which provides a less biased adjusted overall effect. The moderator here is simply the inverse of the 'effective sample size' (see equation 28 in Nakagawa et al. 2022). The slope of this meta-regression provides information about whether funnel plot asymmetry exists and the intercept corresponds to a less biased overall effect size adjusted by small-study effects.

@@

```{r}
dataset_SMDH<-dataset_SMDH%>%
  filter(year_publication!="Unpublished")%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])

MR_SMDH_decline_eff <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = SMDH_variance,# specify lnRR's sampling variance;
                    mods = ~ year_MeanCenter,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

summary(MR_SMDH_decline_eff)

round(r2_ml(MR_SMDH_decline_eff)*100,2)

MR_SMDH_decline_effect_bubble<-orchaRd::mod_results(MR_SMDH_decline_eff,
                     mod="year_MeanCenter",
                     group="paper_ID",
                     weights="prop")
orchaRd::bubble_plot(MR_SMDH_decline_effect_bubble,
                      mod="year_MeanCenter",
                     group="paper_ID")
```

@@

>>>>

# selecting only published dataset
dataset_lnRR_published<-dataset_lnRR%>%
  filter(!is.na(year_publication))

@@

# selecting only published dataset
dataset_SMDH_published<-dataset_SMDH%>%
  filter(!is.na(year_publication))

@@

### Sensitivity Analysis with published data only

VCV_lnRR_published <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_published<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_published)

# saving the model
saveRDS(SA_lnRR_published, file = here::here("sensitivity_analysis/SA_lnRR_published.rds"))

summary(SA_lnRR_published)
predict(SA_lnRR_published)

VCV_SMDH_published <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_SMDH_published<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_published)

# saving the model
saveRDS(SA_SMDH_published, file = here::here("sensitivity_analysis/SA_SMDH_published.rds"))

summary(SA_SMDH_published)
predict(SA_SMDH_published)

@@

```{r Types of parasite}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_lnRR_parasite <- dataset_lnRR%>%
  filter(!is.na(parasite_type))%>%
  mutate(parasite_type=as.factor(parasite_type),
         paper_ID=as.factor(paper_ID),
         Observation_ID=as.factor(Observation_ID),
         group_ID_coded=as.factor(group_ID_coded),
         bird_species=as.factor(bird_species))

@@

@@

Small study effects:

dataset_lnRR_published<-dataset_lnRR_published%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))

lnRR_small_study <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published,# specify lnRR's sampling variance;
                    mods = ~sqrt_effective_n_inv,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_published)

# saving the model
saveRDS(lnRR_small_study, file = here::here("model/lnRR_small_study.rds"))

summary(lnRR_small_study)

round(r2_ml(lnRR_small_study)*100,2)

## Now for SMDH model

dataset_SMDH_published<-dataset_SMDH_published%>%
  mutate(effective_n_inv=1/((4*effective_n_experiment*effective_n_control)/(effective_n_experiment+effective_n_control)))%>%
  mutate(sqrt_effective_n_inv=sqrt(effective_n_inv))

SMDH_small_study <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published,# specify lnRR's sampling variance;
                    mods = ~sqrt_effective_n_inv,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_published)

# saving the model
saveRDS(SMDH_small_study, file = here::here("model/SMDH_small_study.rds"))

summary(SMDH_small_study)

round(r2_ml(SMDH_small_study)*100,2)

Decline effects (Time lag bias)

@@

```{r}
dataset_lnRR_published<-dataset_lnRR_published%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])

lnRR_decline_eff <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published,# specify lnRR's sampling variance;
                    mods = ~ year_MeanCenter,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_published)

# saving the model
saveRDS(lnRR_decline_eff, file = here::here("model/lnRR_decline_eff.rds"))

summary(lnRR_decline_eff)

round(r2_ml(lnRR_decline_eff)*100,2)

## --- Now for SMDH -----
dataset_SMDH_published<-dataset_SMDH_published%>%
  mutate(year_MeanCenter=scale(as.numeric(year_publication),scale=F)[,1])

SMDH_decline_eff <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published,# specify lnRR's sampling variance;
                    mods = ~ year_MeanCenter,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_published)
# saving the model
saveRDS(SMDH_decline_eff, file = here::here("model/SMDH_decline_eff.rds"))

summary(SMDH_decline_eff)

round(r2_ml(SMDH_decline_eff)*100,2)

@@

## All-in Model for heterogeneity

```{r}
## For lnRR complete dataset
VCV_lnRR_published<- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

all_in_lnRR <- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ - 1 + time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      comparision_type + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_published)

## For SMDH complete published dataset
VCV_SMDH_published<- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

all_in_SMDH<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ - 1 + time_of_gnm_addition + 
                      year_MeanCenter + 
                      sqrt_effective_n_inv + 
                      comparision_type + 
                      bird_species + 
                      trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_published)
```

### 2025-06-17 docs/notebook/figures.qmd
Some sensitivity analysis added

@@

```{r orchard plot for intercept only models}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(here,tidyverse,metafor,orchaRd,patchwork,cowplot)

@@

# Plot for lnRR
intercept_lnRR_plot <- orchard_plot(intercept_lnRR,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Log Response Ratio- lnRR)",
                                    colour = "#218833",
                                    fill = TRUE) +
  scale_fill_manual(values = "#218833") +
  scale_colour_manual(values = "#218833") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot <- orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Standardized Mean Difference- SMDH)",
                                    colour = "#895129",
                                    fill = TRUE) +
  scale_fill_manual(values = "#895129") +
  scale_colour_manual(values = "#895129") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

>>>>

# Visualising location and bird species

I am using the figure code from here and adapting it for my needs..

Mentesana, L., Hau, M., D'Amelio, P.B., Adreani, N.M. and Sánchez-Tójar, A. (2025), Do Egg Hormones Have Fitness Consequences in Wild Birds? A Systematic Review and Meta-Analysis. Ecology Letters, 28: e70100. <https://doi.org/10.1111/ele.70100> \[[Code](https://github.com/ASanchez-Tojar/meta-analysis_egg_hormones_and_fitness/blob/main/code/003_figures.R)\]

```{r}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(ggplot2,dplyr,stringr,readr,cowplot)

# Loading cleaned dataset
dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))

# To split population_ID into location and country
dataset_analysis[c('Location','Country')] <- str_split_fixed(dataset_analysis$population_ID, ',', 2)

### For the purpose of this map, I am adding the the longitude and latitude of the study area manually. They exist in the dataset_analysis$population_location as they were extracted from the paper. I have also added some by just searching the general location since the coordinates were not provided for these..

data_coordinates<-data.frame(Location = unique(dataset_analysis$Location),
                             Lat_Location=c(51.50,38.41,40.45,40.75,45.40,52.39,37.18,
                                             40.53,54,39.05,
                                             51.57,48,42.54,39.95,
                                             40.13,47.96,28.10,52.06,60.39),
                             Long_Location=c(19.29,-9.02,-3.50,-3.78,-64.10,6.05,-3.11,
                                              -4.0,-124,-76.81,
                                              -2.26,11,8.91,-75.16,
                                              -8.27,11.18,107.10,8.25,5.32))

## Let's now merge the datasets by the location names

dataset_analysis <- dataset_analysis %>%
  left_join(data_coordinates, by = "Location")

## Count how many times each Location appears in the full dataset
location_counts <- dataset_analysis %>%
  group_by(Location) %>%
  summarise(size.freq = n_distinct(paper_ID), .groups = 'drop')

# I will subset the data base so that I get one row per Location AND bird_species.
# Like this, I get one point per location independently of the number of effect sizes 
# we have per place.
dataset_analysis_unique_rows <- dataset_analysis %>%
  distinct(Location, bird_species, .keep_all = TRUE)

# Merge the counts into the unique rows dataset
dataset_analysis_unique_rows <- dataset_analysis_unique_rows %>%
  left_join(location_counts, by = "Location")

# create data for world coordinates using  
# map_data() function 
world_coordinates <- map_data("world")

# create world map using ggplot() function 
# geom_map() function takes world coordinates as input to plot world map 
fig_map <- ggplot() + 
  geom_map(data = world_coordinates, map = world_coordinates, 
    aes(long, lat, map_id = region), 
    color = "white", fill= "#D9EAF7") + 
  geom_point(data = dataset_analysis_unique_rows, 
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour=bird_species), alpha = 0.6) +
scale_size(range = c(3, 10)) +
  scale_fill_manual(values = c("#11a3d9","#f6c800",  "#fb5158", 
                               "#ea60e9", "#40ad28","#ef8c27","#a083e2")) +

  scale_colour_manual(values = c("#11a3d9","#f6c800",  "#fb5158", 
                               "#ea60e9", "#40ad28","#ef8c27","#a083e2")) +
  guides(size = "none", alpha = "none") +
  labs(x = "Longitude",
       y = "Latitude")+
  theme_classic() +
  theme(legend.position = c(0.88, 0.10),
        legend.justification = c("right", "bottom"),
        legend.background = element_rect(fill = "white", color = "black"),
        axis.title = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_blank())

# To add the zoomed in map.. to show the points more clearly
zoom_map <- ggplot() + 
  geom_map(data = world_coordinates, map = world_coordinates,
           aes(long, lat, map_id = region),
           color = "white", fill = "lightgrey") +
  geom_point(data = dataset_analysis_unique_rows,
             aes(x = Long_Location, y = Lat_Location,
                 size = size.freq, colour = bird_species),
             alpha = 0.6) +
  scale_size(range = c(3, 10)) +
  scale_colour_manual(values = c("#11a3d9","#f6c800","#fb5158",
                                 "#ea60e9","#40ad28","#ef8c27","#a083e2")) +
  coord_cartesian(xlim = c(-120, 120), ylim = c(25, 65)) +  # Northern Hemisphere
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(color = "white", fill="white"),
    panel.border = element_rect(colour = "black", fill = NA, size = 1)
  )

geographic_distribution <- ggdraw() +
  draw_plot(fig_map) +
  draw_plot(zoom_map, x = 0.10, y = 0.12, width = 0.62, height = 0.22)

ggsave(here::here("figures/Figure_Distribution.png"), plot = geographic_distribution, height = 250, width = 400, units = "mm", dpi = 600)
```

# Visualizing author network

```{r}
# # Loading cleaned dataset
# dataset_analysis<- read_csv(here::here("data/03_data_cleaning/dataset_analysis.csv"))
# 
# library(tidyverse)
# library(tidyr)
# library(igraph)
# installed.packages("ggraph")
# library(ggraph)
# 
# clean_author <- function(name) {
#   name %>%
#     str_replace_all(",", "") %>%   # remove commas
#     str_squish() %>%              # remove excess whitespace
#     str_to_title()                # standardize case (optional, if needed)
# }
# 
# # Apply cleaning to first and last author fields
# author_roles <- dataset_analysis %>%
#   distinct(paper_ID, .keep_all = TRUE) %>%
#   select(paper_ID, authors, bird_species) %>%
#   mutate(
#     author_list = str_split(authors, ";\\s*"),
#     first_author = map_chr(author_list, 1),
#     last_author = map_chr(author_list, ~ tail(.x, 1))
#   )%>%
#   mutate(
#     first_author=case_when(
#       first_author == "Polo, V" | first_author == "Polo V" ~ "Polo V.",
#       first_author == "Soler J J" ~ "Soler J. J.",
#       first_author == "Glądalski, M" ~ "Gladalski M.",
#       first_author == "Gwinner, H" | first_author == "Gwinner H" ~ "Gwinner H.",
#       TRUE ~ as.character(first_author)
#     ))%>%
#   mutate(
#     last_author=case_when(
#       last_author == "Veiga J P" ~ "Veiga J. P.",
#       TRUE ~ as.character(last_author)
#     )
#   )
# 
# # 2. Create Core Author Pairs (First → Last)
# author_pairs_core <- author_roles %>%
#   select(paper_ID, first_author, last_author) %>%
#   rename(from = first_author, to = last_author) %>%
#   filter(from != to) %>%
#   count(from, to, name = "weight")
# 
# # 3. Build Author Graph
# author_graph <- graph_from_data_frame(author_pairs_core, directed = TRUE)
# 
# # 4. Assign Dominant Bird Species to PIs (last authors)
# author_species <- author_roles %>%
#   select(last_author, bird_species) %>%
#   group_by(last_author, bird_species) %>%
#   count() %>%
#   slice_max(n, n = 1) %>%
#   distinct(last_author, .keep_all = TRUE)
# 
# # Add species as vertex attribute
# V(author_graph)$bird_species <- author_species$bird_species[
#   match(V(author_graph)$name, author_species$last_author)
# ]
# 
# # 5. Optional: Scale node size by number of papers per author
# author_paper_counts <- author_roles %>%
#   pivot_longer(cols = c(first_author, last_author), names_to = "role", values_to = "author") %>%
#   count(author, name = "paper_count")
# 
# # Add paper count attribute
# V(author_graph)$paper_count <- author_paper_counts$paper_count[
#   match(V(author_graph)$name, author_paper_counts$author)
# ]
# 
# # Replace NAs with small size for isolated authors
# V(author_graph)$paper_count[is.na(V(author_graph)$paper_count)] <- 1
# 
# # 6. Plot with ggraph
# ggraph(author_graph, layout = "fr") +
#   geom_edge_link(aes(width = weight), alpha = 0.4, color = "grey50") +
#   geom_node_point(aes(color = bird_species, size = paper_count)) +
#   geom_node_text(aes(label = name), repel = TRUE, size = 3) +
#   scale_edge_width(range = c(0.3, 2)) +
#   scale_size(range = c(3, 10)) +
#   theme_void() +
#   labs(title = "Author Collaboration Network by Bird Species",
#        size = "Paper Count") +
#   theme(legend.position = "right")
#

#

@@

```{r orchard plot for intercept only models}
pacman::p_load(here,tidyverse,metafor,orchaRd,patchwork,cowplot)

@@

# Plot for lnRR
intercept_lnRR_plot <- orchaRd::orchard_plot(intercept_lnRR,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Log Response Ratio- lnRR)",
                                    fill = T) +
   scale_fill_manual(values = "#218833") +
   scale_colour_manual(values = "#218833" ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0),
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")

# Plot for SMDH
intercept_SMDH_plot <- orchaRd::orchard_plot(intercept_SMDH,
                                    mod = "1",
                                    group = "paper_ID",
                                    trunk.size = 0.5,
                                    branch.size = 2,
                                    twig.size = 0.5,
                                    alpha = 0.3,
                                    xlab = "Effect size (Standardized Mean Difference- SMDH)",
                                    fill = TRUE) +
  scale_fill_manual(values = "#895129") +
  scale_colour_manual(values = "#895129") +
  theme_minimal(base_size = 14) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    legend.position = c(0.98, 0.05),
    legend.justification = c(1, 0), 
    legend.direction = "horizontal",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_blank()
  ) +
  labs(x = "Intercept")
